---
title: "02_assessment (IN PROGRESS)"
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
---

The majority of the work has been the evaluation of the papers, the consolidation of the assessments, and the still-to-do **synthesis of the rich feedback** as well as interpretation of our findings. 


- When a paper has `NA` in the three columns `consolidated_data`, `consolidated_methods` and `consolidated_results`, this mans it is a conceptual paper (`consolidated_cp` is `TRUE`).
- still-to-do synthesis of the rich feedback as well as interpretation of our findings.


```{r}
#| label: load-packages
#| message: false

library(here)
library(tidyverse)
library(scales)
```

```{r}
#| label: load-data
#| message: false

assessment <- readr::read_csv(here::here("data-clean", "assessment.csv"))

# TODO: forcats: "udao_" columns


```

Some basic facts & checks:

```{r}
#| label: basic-indicators
#| message: false

n_papers <- nrow(assessment)


n_cp <- nrow(filter(assessment, consolidated_cp == TRUE))
p_cp <- n_cp / n_papers


```

- Total number of assessed papers: `{r} n_papers`
- Number of `CP` papers: `{r} n_cp` (`{r} scales::label_percent(accuracy = 0.1)(p_cp)`)
- Number of eligible papers for analysis: `{r} n_papers - n_cp` (`{r} scales::label_percent(accuracy = 0.1)(1 - p_cp)`)