---
title: "08_authorship"
description: |
    Recreates the pre- vs post-intervention reproducibility-level distribution figure
    from `03_results_reprolevels.qmd` (manuscript Figures 2 / 3, a.k.a. `fig:agile` and
    `fig:giscience` in `manuscript/main.tex`; the JOSIS PDF cites the source as
    `04_results_reprolevels.qmd`, the file's earlier name), but **pooling both
    conferences** and restricting to papers with **at least one cross-conference
    author** -- i.e. papers that have at least one author whose resolved identity also
    appears on a paper at the other conference series in the corpus.

    Cross-author identities are built in `07_authorship.ipynb` (section 7, two-pass
    identity merge with ORCID-first and normalised-name fallback, plus any manual
    resolutions in `data/orcid-resolutions.csv`) and written to `data-clean/authors.csv`.
    This notebook consumes that file; re-run `07_authorship.ipynb` first if the
    identity merge has changed.
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
---

```{r}
#| label: load-packages
#| message: false

library(here)
library(tidyverse)
library(dplyr)
library(scales)
library(stringr)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(RColorBrewer)
```

## Helpers

Copied verbatim from `03_results_reprolevels.qmd` so this notebook is self-contained
and the panels look identical to the manuscript figures.

```{r}
#| label: functions

criteriaBarplot2 <- function(df, col, main, colour,
                             agg_level = FALSE, agg_level_y = 35) {
  col <- enquo(col)
  p <-
    ggplot2::ggplot(df, aes(x = !!col, y = n)) +
    ggplot2::geom_col(fill = colour, color = "black") +
    ggplot2::geom_text(
      aes(label = scales::label_percent(accuracy = 0.1)(p)),
      vjust = -0.5,
      size = 3.5,
      color = "black"
    ) +
    ggplot2::geom_vline(xintercept = 2.5,
                        linetype = "dotted",
                        color = "gray",
                        linewidth = 1) +
    ggplot2::ggtitle(main) +
    ggplot2::xlab("Level") +
    ggplot2::xlim(c("U", "D", "A", "O")) +
    ggplot2::ylab("") +
    ggplot2::scale_y_continuous(breaks = breaks,
                                limits = range(breaks)) +
    ggthemes::theme_tufte(base_size = 10) +
    ggplot2::theme(axis.ticks.x = element_blank())

  if (agg_level) {
    p <- p +
      ggplot2::annotate("text", x = 2:3, y = agg_level_y,
                        label = c("LOW", "HIGH"), alpha = 0.5)
  }
  return(p)
}

count_pct <- function(df, col) {
  col <- enquo(col)
  df |>
    dplyr::count(!!col) |>
    dplyr::mutate(p = n / sum(n))
}
```

## Load corpus and cross-author subset

```{r}
#| label: load-data
#| message: false

paper_evaluation <-
  readr::read_csv(here::here("data-clean", "all-data.csv"),
                  na = c("NA", "Not applicable")) |>
  dplyr::filter(consolidated_cp == FALSE)

authors <- readr::read_csv(
  here::here("data-clean", "authors.csv"),
  na = c(""),
  show_col_types = FALSE
)

# Cross-conference identities: resolved identities that appear in BOTH conference
# series (any year, direction pooled -- matches section 8 of 07_authorship.ipynb).
agile_ids <- authors |>
  dplyr::filter(conf == "agile") |>
  dplyr::pull(identity) |>
  unique()
giscience_ids <- authors |>
  dplyr::filter(conf == "giscience") |>
  dplyr::pull(identity) |>
  unique()
cross_ids <- intersect(agile_ids, giscience_ids)

# Papers (from either conference) with >= 1 cross-conference author.
cross_papers <- authors |>
  dplyr::filter(identity %in% cross_ids) |>
  dplyr::pull(paper) |>
  unique()

paper_evaluation_cross <- paper_evaluation |>
  dplyr::filter(paper %in% cross_papers)

message(stringr::str_glue(
  "{length(cross_ids)} cross-conference author identities; ",
  "{length(cross_papers)} papers with at least one such author ",
  "(of {nrow(paper_evaluation)} non-conceptual papers)."
))
```

## Cross-author papers -- pre- vs post-intervention (both conferences pooled)

The intervention split mirrors `03_results_reprolevels.qmd`: *pre* = `year < 2020`,
*post* = `year > 2020`, with 2020 excluded as the AGILE transition year (the AGILE
Reproducibility Guidelines were optional that year). Both AGILE and GIScience papers
are pooled -- the question this figure answers is how the *set of cross-author papers
as a whole* distributes across U / D / A / O for each criterion before and after the
intervention.

```{r}
#| label: cross-subset

papers_pre_cross  <- paper_evaluation_cross |> dplyr::filter(year <  2020)
papers_post_cross <- paper_evaluation_cross |> dplyr::filter(year >  2020)

n_pre  <- nrow(papers_pre_cross)
n_post <- nrow(papers_post_cross)

message(stringr::str_glue(
  "Cross-author papers (both conferences pooled): pre={n_pre}, post={n_post}."
))

# Breakdown per conference, for transparency in the text.
pre_by_conf  <- papers_pre_cross  |> dplyr::count(conf)
post_by_conf <- papers_post_cross |> dplyr::count(conf)
message("Pre by conference:")
print(pre_by_conf)
message("Post by conference:")
print(post_by_conf)

d_data_pre     <- count_pct(papers_pre_cross,  consolidated_data)
d_methods_pre  <- count_pct(papers_pre_cross,  consolidated_methods)
d_results_pre  <- count_pct(papers_pre_cross,  consolidated_results)
d_data_post    <- count_pct(papers_post_cross, consolidated_data)
d_methods_post <- count_pct(papers_post_cross, consolidated_methods)
d_results_post <- count_pct(papers_post_cross, consolidated_results)
```

```{r}
#| label: barplots
#| warning: false

colours <- RColorBrewer::brewer.pal(3, "Set1")

# Pooled cross-author subset: ~29 pre, ~38 post papers. Max bin can approach ~20.
# The y-axis is fixed to 0--35 so this figure is visually comparable to the GIScience
# pre-vs-post figure in 03_results_reprolevels.qmd, which uses the same scale.
breaks <- seq(from = 0, to = 35, by = 5)

plotA <- criteriaBarplot2(d_data_pre,
                          col = consolidated_data,
                          main = "Input data",
                          colour = colours[1],
                          agg_level = TRUE, agg_level_y = 35)
plotB <- criteriaBarplot2(d_methods_pre,
                          col = consolidated_methods,
                          main = "Methods, analysis, processing",
                          colour = colours[2])
plotC <- criteriaBarplot2(d_results_pre,
                          col = consolidated_results,
                          main = "Results",
                          colour = colours[3])

plotD <- criteriaBarplot2(d_data_post,
                          col = consolidated_data,
                          main = "Input data",
                          colour = colours[1],
                          agg_level = TRUE, agg_level_y = 35)
plotE <- criteriaBarplot2(d_methods_post,
                          col = consolidated_methods,
                          main = "Methods, analysis, processing",
                          colour = colours[2])
plotF <- criteriaBarplot2(d_results_post,
                          col = consolidated_results,
                          main = "Results",
                          colour = colours[3])
```

```{r}
#| label: fig-cross-author-pre-post
#| warning: false
#| fig-cap: "Evaluation of the potential reproducibility level of cross-conference-author papers (both AGILE and GIScience pooled) pre- vs post-intervention, for each criterion: Data (A, D), Methods (B, E) and Results (C, F)."
#| fig-width: 10
#| fig-asp: 0.8
#| out-width: 100%
#| fig-align: center

(plotA + plotB + plotC) / (plotD + plotE + plotF) +
  plot_layout(widths = c(1, 1, 1)) +
  plot_annotation(
    title = "Cross-conference-author papers (both conferences pooled), assessed per criterion.",
    subtitle = stringr::str_glue(
      "Top: pre-intervention (year<2020) papers (N={n_pre}). ",
      "Bottom: post-intervention (year>2020) papers (N={n_post}).\n ",
      "X-axis labels: (U)ndocumented; (D)ocumented; Documented and (A)vailable; ",
      "Documented, Available and (O)pen."
    ),
    caption = stringr::str_glue(
      "2020 excluded (AGILE transition year). A paper counts as cross-author if at ",
      "least one of its authors (resolved identity, pooling all years) also appears ",
      "on a paper at the other conference. Cross-conference authors in the corpus: ",
      "{length(cross_ids)}."
    ),
    tag_levels = "A"
  ) &
  theme(plot.tag = element_text(size = 8))
```

```{r}
#| label: fig-cross-author-pre-post-save
#| warning: false
#| echo: false

ggplot2::ggsave(
  file = here::here("figs", "cross_author_pre_post.png"),
  width = 10, height = 8, dpi = 300
)
```
